\section{Introduction}

This codelab will help you build and run a simple chat application using Dart. Along the way, you will learn:

\begin{itemize}
\item The fundamentals of the Dart programming language
\item How to use the Dart HTML libraries
\item Bi-directional communication with WebSockets
\item The basics of Dart Editor
\item How to find answers from the online resources
\end{itemize}

\subsection{Prerequisites}

This codelab assumes you are familiar with fundamental web programming. You should know the basics of HTML, CSS, and JavaScript. You should also have experience with a programming language such as Java, C\#, or JavaScript.

This codelab assumes you have watched the following video tutorials:

\begin{itemize}
\item Google I/O 101: Introduction to Dart with Seth Ladd
\item Google I/O 101: Dart Editor with Devon Carew
\end{itemize}

\subsection{Installs}

This codelab requires the Dart Editor. You can find builds of the editor for Mac, Windows, and Linux at \url{http://dartlang.org/editor/}.

\subsection{Additional materials}

This codelab provides easy to follow, step-by-step instructions. However, you’ll find it quite useful to have the following online resources loaded up and ready to access.

\begin{itemize}
\item \url{http://api.dartlang.org} The Dart API docs list the functions, classes, and methods for all the libraries.
\item \url{http://www.dartlang.org/docs/language-tour/} The Dart Language Tour is your high-level guide through the Dart language.
\item \url{http://www.dartlang.org/docs/library-tour/} The Dart Library Tour covers most of the bundled libraries of the Dart platform.
\item \url{http://synonym.dartlang.org/} Know JavaScript? This is your resource! Translate common idioms, patterns, and snippets from JavaScript to Dart.
\end{itemize}
